---
title: "Pre Practice ISLR"
author: "Diana McSpadden"
date: "1/11/2021"
output: html_document
---

```{r}
x=rnorm (50)
y=x+rnorm (50, mean=50, sd=.1)
print(cor(x,y))
```


```{r}
plot(x,y,xlab=" this is the x-axis",ylab=" this is the y-axis",
main=" Plot of X vs Y")
```


```{r}
x=seq(-pi ,pi ,length =50)
y=x
f=outer(x,y,function (x,y)cos(y)/(1+x^2))
contour (x,y,f)
contour (x,y,f,nlevels =45, add=T)
fa=(f-t(f))/2
contour (x,y,fa,nlevels =15)
```


```{r}
image(x,y,fa)
persp(x,y,fa)
persp(x,y,fa ,theta =30)
persp(x,y,fa ,theta =30, phi =20)
persp(x,y,fa ,theta =30, phi =70)
persp(x,y,fa ,theta =30, phi =40)
```


```{r}
A <- matrix (1:16 ,4 ,4)
A
```


```{r}
A[2,3]
```


```{r}
A[c(1,3) ,c(2,4) ]
```


```{r}
A[-c(1,3) ,]
```


```{r}
dim(A)
```


```{r}
dim(A)[2]
```
1. For each of parts (a) through (d), indicate whether we would generally
expect the performance of a flexible statistical learning method to be
better or worse than an inflexible method. Justify your answer.

(a) The sample size n is extremely large, and the number of predictors p is small.
ANSWER: inflexible

(b) The number of predictors p is extremely large, and the number of observations n is small.
ANSWER: inflexible

(c) The relationship between the predictors and response is highly non-linear.
ANSWER: flexible

(d) The variance of the error terms, i.e. σ2 = Var(E), is extremely high.
ANSWER: flexible


2. Explain whether each scenario is a classification or regression problem,and indicate whether we are most interested in inference or prediction.

Finally, provide n and p.

(a) We collect a set of data on the top 500 firms in the US. For each
firm we record profit, number of employees, industry and the
CEO salary. We are interested in understanding which factors
affect CEO salary.
ANSWER: n = 500, p = 3, regression

(b) We are considering launching a new product and wish to know
whether it will be a success or a failure. We collect data on 20
similar products that were previously launched. For each product
we have recorded whether it was a success or failure, price
charged for the product, marketing budget, competition price,
and ten other variables.
ANSWER: categorization, n = 20, p = 13

(c) We are interest in predicting the % change in the USD/Euro
ANSWER: regression

## Applied Chapter 2

```{r}
college <- read.csv("https://statlearning.com/College.csv")
head(college)
```
(b) Look at the data using the fix() function. You should notice
that the first column is just the name of each university.We don’t
really want R to treat this as data. However, it may be handy to
have these names for later. Try the following commands:

```{r}
rownames(college) <- college[, 1]
college <- college[, -1]
```


```{r}
fix (college )
```
```{r}
summary(college)
```


```{r}
pairs(college[,2:11])
```

```{r}
attach(college)
# plot side by side boxplots of Outstate versus Private (Private universities have more out of state students):
boxplot(college$Outstate~college$Private)
```
Create a new qualitative variable, called Elite, by binning the Top10perc variable. We are going to divide universities into two groups based on whether or not the proportion
of students coming from the top 10% of their high school classes exceeds 50%.


```{r}
Elite <- rep("No", nrow(college))
Elite[college$Top10perc > 50] <- "Yes"

college <- data.frame(college, Elite)

head(college)
```


```{r}
boxplot(college$Outstate~college$Elite)

```
v. Use the hist() function to produce some histograms with differing numbers of bins for a few of the quantitative variables. You may find the command par(mfrow=c(2,2)) useful: it will divide the print window into four regions so that four plots can be made simultaneously. Modifying the arguments to this function will divide the screen in other ways.

```{r}
attach(college)
par(mfrow=c(2,2))

hist(PhD, breaks=3, main="With breaks=3")

hist(Expend, breaks=5, main="With breaks=5")

hist(Top10perc, breaks=10, main="With breaks=10")

hist(Grad.Rate, breaks=3, main="With breaks=3")

```















