---
title: 'Haiti: Load Hold-out Data'
author: "Diana McSpadden"
date: "4/16/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```


```{r}
# directory
holdout_dir = "HoldOut"

# unzip
unzip(file.path(holdout_dir,"Hold+Out+Data.zip"), exdir = holdout_dir)
```

### Not Blue Tarps

File 1
<ul>
<li>Name: orthovnir057_ROI_NON_Blue_Tarps.txt</li>
<li>**NOT** Blue Tarp</li>
</ul>

```{r notBlueFile1-1}

read_lines(file.path(holdout_dir, "orthovnir057_ROI_NON_Blue_Tarps.txt"), n_max=12)

```
<ul>
<li>8th row is header</li>
<li>B1 == Red</li>
<li>B2 == Blue</li>
<li>B3 == Green</li>
</ul>

```{r notBlueFile1-2}

df_NotBlue_057 = as.data.frame(read_table(file.path(holdout_dir, "orthovnir057_ROI_NON_Blue_Tarps.txt"), 
                               skip=7, col_types = cols(';'="-", ID = "c"))) %>%
                              rename(Red = B1, Green = B2, Blue = B3) %>%
                              mutate(ClassBinary = '0') %>%
                              select(Lat, Lon, Red, Green, Blue, ClassBinary)

head(df_NotBlue_057)

dim(df_NotBlue_057)
```


File 2
<ul>
<li>Name: orthovnir067_ROI_NOT_Blue_Tarps.txt</li>
<li>**NOT** Blue Tarp</li>
</ul>

```{r notBlueFile2-1}

read_lines(file.path(holdout_dir, "orthovnir067_ROI_NOT_Blue_Tarps.txt"), n_max=12)

```

```{r notBlueFile2-2}

df_NotBlue_067 = as.data.frame(read_table(file.path(holdout_dir, "orthovnir067_ROI_NOT_Blue_Tarps.txt"), 
                               skip=7, col_types = cols(';'="-", ID = "c"))) %>%
                              rename(Red = B1, Green = B2, Blue = B3) %>%
                              mutate(ClassBinary = '0') %>%
                              select(Lat, Lon, Red, Green, Blue, ClassBinary)

head(df_NotBlue_067)

dim(df_NotBlue_067)
```

File 3
<ul>
<li>Name: orthovnir069_ROI_NOT_Blue_Tarps.txt</li>
<li>**NOT** Blue Tarp</li>
</ul>


```{r notBlueFile3-1}

read_lines(file.path(holdout_dir, "orthovnir069_ROI_NOT_Blue_Tarps.txt"), n_max=12)

```
```{r notBlueFile3-2}

df_NotBlue_069 = as.data.frame(read_table(file.path(holdout_dir, "orthovnir069_ROI_NOT_Blue_Tarps.txt"), 
                               skip=7, col_types = cols(';'="-", ID = "c"))) %>%
                              rename(Red = B1, Green = B2, Blue = B3) %>%
                              mutate(ClassBinary = '0') %>%
                              select(Lat, Lon, Red, Green, Blue, ClassBinary)

head(df_NotBlue_069)

dim(df_NotBlue_069)
```


File 4
<ul>
<li>Name: orthovnir078_ROI_NON_Blue_Tarps.txt</li>
<li>**NOT** Blue Tarp</li>
</ul>

```{r notBlueFile4-1}

read_lines(file.path(holdout_dir, "orthovnir078_ROI_NON_Blue_Tarps.txt"), n_max=12)

```

```{r notBlueFile4-2}

df_NotBlue_078 = as.data.frame(read_table(file.path(holdout_dir, "orthovnir078_ROI_NON_Blue_Tarps.txt"), 
                               skip=7, col_types = cols(';'="-", ID = "c"))) %>%
                              rename(Red = B1, Green = B2, Blue = B3) %>%
                              mutate(ClassBinary = '0') %>%
                              select(Lat, Lon, Red, Green, Blue, ClassBinary)

head(df_NotBlue_078)

dim(df_NotBlue_078)
```

### Blue Tarps

File 1
<ul>
<li>Name: orthovnir067_ROI_Blue_Tarps.txt</li>
<li>**Blue Tarp**</li>
</ul>

```{r BlueFile1-1}

read_lines(file.path(holdout_dir, "orthovnir067_ROI_Blue_Tarps.txt"), n_max=12)

```
```{r BlueFile1-2}

df_Blue_067 = as.data.frame(read_table(file.path(holdout_dir, "orthovnir067_ROI_Blue_Tarps.txt"), 
                               skip=7, col_types = cols(';'="-", ID = "c"))) %>%
                              rename(Red = B1, Green = B2, Blue = B3) %>%
                              mutate(ClassBinary = '1') %>%
                              select(Lat, Lon, Red, Green, Blue, ClassBinary)

head(df_Blue_067)

dim(df_Blue_067)
```

File 2
<ul>
<li>Name: orthovnir069_ROI_Blue_Tarps.txt</li>
<li>**Blue Tarp**</li>
</ul>

```{r BlueFile2-1}

read_lines(file.path(holdout_dir, "orthovnir069_ROI_Blue_Tarps.txt"), n_max=12)

```


```{r BlueFile2-2}

df_Blue_069 = as.data.frame(read_table(file.path(holdout_dir, "orthovnir069_ROI_Blue_Tarps.txt"), 
                               skip=7, col_types = cols(';'="-", ID = "c"))) %>%
                              rename(Red = B1, Green = B2, Blue = B3) %>%
                              mutate(ClassBinary = '1') %>%
                              select(Lat, Lon, Red, Green, Blue, ClassBinary)

head(df_Blue_069)

dim(df_Blue_069)
```

File 3
<ul>
<li>Name: orthovnir078_ROI_Blue_Tarps.txt</li>
<li>**Blue Tarp**</li>
</ul>

```{r BlueFile3-1}

read_lines(file.path(holdout_dir, "orthovnir078_ROI_Blue_Tarps.txt"), n_max=12)

```

```{r BlueFile3-2}
df_Blue_078 = as.data.frame(read_table(file.path(holdout_dir, "orthovnir078_ROI_Blue_Tarps.txt"), 
                               skip=7, col_types = cols(';'="-", ID = "c"))) %>%
                              rename(Red = B1, Green = B2, Blue = B3) %>%
                              mutate(ClassBinary = '1') %>%
                              select(Lat, Lon, Red, Green, Blue, ClassBinary)

head(df_Blue_078)

dim(df_Blue_078)
```


```{r JoinDFs}
# join the 7 dataframes of hold out data
df_HoldOut = dplyr::bind_rows(df_NotBlue_057, df_NotBlue_067, df_NotBlue_069, df_NotBlue_078, df_Blue_067, df_Blue_069, df_Blue_078)

dim(df_HoldOut)
head(df_HoldOut)

```

```{r saveDF}
library(readr)

write_csv(df_HoldOut, "holdout.csv")

```



