---
title: "Module 4 HW"
author: "Diana McSpadden"
date: "2/28/2021"
output: html_document
---

# HW Module 4
## H. Diana McSpadden (hdm5s)
## 2021-03-01


You will complete exercises #3, 5, 6, 7, and 8 at the end of chapter 5 (p. 197-201) of your textbook.

# Q3. We now review k-fold cross-validation.
## (a) Explain how k-fold cross-validation is implemented.
<br /><br />
**Answer Q3a:** k-fold cross-validation approach divides the data into k number of groups. One-by-one each of the groups is used as the test/validation data, while the remaining other k-1 groups are used as the training data. The model is trained on the k-1 groups, tested on the hold out group, and the mean error rate for the k-folds is used to evaluate the model.

## (b) What are the advantages and disadvantages of k-fold cross-validation relative to:
<br /><br />
**Answer Q3b i and ii:**
<br />
An advantage in comparison to LOOCV is that fewer iterations need to be run. In LOOCV n models are fit and tested. In k-fold, only k models are fit and tested, and typically many fewer than n models are used.
<br /><br />
k-fold cross validation also provides better estimates of error because it has less bias than a 50-50 split, but a littel more bias than a LOOCV approach because more data is used to fit the model than in an 50-50 split, but less data than in LOOCV. However, variance is much improved because the average error of k models, vs. n models allows for more variance to be accounted for as the training data sets are more difference from each other when using the k-fold method.


# Q5
In Chapter 4, we used logistic regression to predict the probability of default using income and balance on the Default data set. We will now estimate the test error of this logistic regression model using the validation set approach. Do not forget to set a random seed before beginning your analysis.

## (a) Fit a logistic regression model that uses income and balance to predict default.
```{r q5a-1, message=FALSE}
library(ISLR)
library(tidyverse)
library(GGally)
library(dplyr)
library(broom)
library(caret)
library(class)
library(purrr)
library(gridExtra)
library(ROCR)
library(yardstick)
```

```{r q5a-2}
head(Default)
```

Create the factor/dummy variable for student:
```{r q5a-3}
df.Default = Default %>%
  mutate(student = factor(student))
```

```{r q5a-4}
head(df.Default)
```

```{r q5a-4}
log.model = glm(default~income+balance, data=df.Default, family=binomial)
```

```{r}
summary(log.model)
```


## (b) Using the validation set approach, ...
estimate the test error of this model. In order to do this, you must perform the following steps:

**i. Split the sample set into a training set and a validation set.**

```{r q5b-1}
set.seed(1976)

# create an 80/20 split
sample_size = floor(0.8*nrow(df.Default))

# randomly split data
picked = sample(seq_len(nrow(df.Default)),size = sample_size)
df.Default.train = df.Default[picked,]
df.Default.test = df.Default[-picked,]

dim(df.Default.train)
dim(df.Default.test)

```


**ii. Fit a multiple logistic regression model using only the training observations.**

```{r q5b-2}

log.model.split = glm(default ~ income+balance, family = binomial, data = df.Default.train)
summary(log.model.split)

```


**iii. Obtain a prediction of default status for each individual in the validation set by computing the posterior probability of default for that individual, and classifying the individual to the default category if the posterior probability is greater than 0.5.**

``` {r q5b-3}
predictions= log.model.split %>% augment(newdata=df.Default.test) %>% 
  dplyr::select(default, .fitted)  %>% 
  mutate(.prediction=ifelse(.fitted<0, "No", "Yes")) %>%
  mutate(.prediction=fct_relevel(as_factor(.prediction), c("No", "Yes"))) 

predictions %>% conf_mat(default, .prediction)

```

**iv. Compute the validation set error, which is the fraction of the observations in the validation set that are misclassified.**

```{r q5b-4}
print(paste("Split Sample Validation Set Error ((42 + 8) / 2000): ", (42+ 8) / 2000))
```
**Answer Q5b iv: ** Split Sample Validation Set Error ((42 + 8) / 2000):  0.025

##(c) Repeat the process in (b) three times, ...
using three different splits of the observations into a training set and a validation set. Comment on the results obtained.

```{r q5c-1}
for (i in seq(1,3))
{
  set.seed(1976 + i)

  # create an 80/20 split
  sample_size = floor(0.8*nrow(df.Default))
  
  # randomly split data
  picked = sample(seq_len(nrow(df.Default)),size = sample_size)
  df.Default.train = df.Default[picked,]
  df.Default.test = df.Default[-picked,]

  
  log.model.split = glm(default ~ income+balance, family = binomial, data = df.Default.train)
  
  
  predictions= log.model.split %>% augment(newdata=df.Default.test) %>% 
    dplyr::select(default, .fitted)  %>% 
    mutate(.prediction=ifelse(.fitted<0, "No", "Yes")) %>%
    mutate(.prediction=fct_relevel(as_factor(.prediction), c("No", "Yes"))) 

  print(paste("Confusion Matrix ", i, ": "))
  
  print(predictions %>% conf_mat(default, .prediction))
  
}
```

**Answer Q5c: ** For the three different samplings produced test set errors of:
<ul>
<li>(46 + 5) / 2000 = 0.0255</li>
<li>(46 + 11) / 2000 = 0.0285</li>
<li>(47 + 6) / 2000 = 0.0265</li>
<li>And our original split produced an error of 0.025</li>
</ul>
<br /><br />
The mean of these errors is **0.0264** which is most likely closer to the true error rate than any of the single train/test splits.

## (d) Now consider a logistic regression model ...
**that predicts the probability of default using income, balance, and a dummy variable for student.** 
<br />
**Estimate the test error for this model using the validation set approach.**

```{r q5d-1}
for (i in seq(0,3))
{
  set.seed(1976 + i)

  # create an 80/20 split
  sample_size = floor(0.8*nrow(df.Default))
  
  # randomly split data
  picked = sample(seq_len(nrow(df.Default)),size = sample_size)
  df.Default.train = df.Default[picked,]
  df.Default.test = df.Default[-picked,]

  
  log.model.split = glm(default ~ income+balance+student, family = binomial, data = df.Default.train)
  
  
  predictions= log.model.split %>% augment(newdata=df.Default.test) %>% 
    dplyr::select(default, .fitted)  %>% 
    mutate(.prediction=ifelse(.fitted<0, "No", "Yes")) %>%
    mutate(.prediction=fct_relevel(as_factor(.prediction), c("No", "Yes"))) 

  print(paste("Confusion Matrix ", i, ": "))
  
  print(predictions %>% conf_mat(default, .prediction))
  
}
```

**Comment on whether or not including a dummy variable for student leads to a reduction in the test error rate.**
<br /><br />
I used the same seeds as were used in the models that did not include the student dummy variable.<br />
My four test/train split repetitions produced the following errors:
<ul>
<li>55 / 2000 = 0.0275</li>
<li>51 / 2000 = 0.0255</li>
<li>58 / 2000 = 0.029</li>
<li>55 / 2000 = 0.0275</li>
</ul>

The mean of these 4 repetitions is 0.0274 which is slightly higher than the models without student. Including student does **not** result in a decreased error rate based on my tests.

# Q6.
We continue to consider the use of a logistic regression model to predict the probability of default using income and balance on the Default data set. In particular, we will now compute estimates for the standard errors of the income and balance logistic regression coefficients in two different ways: 
<ol>
<li> using the bootstrap, and </li>
<li>using the standard formula for computing the standard errors in the glm() function. Do not forget to set a random seed before beginning your analysis.</li>
</ol>

## (a) Using the summary() and glm() functions, ...
determine the estimated standard errors for the coefficients associated with income and balance in a multiple logistic regression model that uses both predictors.


## (b) Write a function, boot.fn(), ...
that takes as input the Default data set as well as an index of the observations, and that outputs the coefficient estimates for income and balance in the multiple logistic regression model.


## (c) Use the boot() function together with your boot.fn() function...
to estimate the standard errors of the logistic regression coefficients for income and balance.


## (d) Comment on the estimated standard errors ...
obtained using the glm() function and using your bootstrap function.


# Q7. 
In Sections 5.3.2 and 5.3.3, we saw that the cv.glm() function can be used in order to compute the LOOCV test error estimate. Alternatively, one could compute those quantities using just the glm() and predict.glm() functions, and a for loop. You will now take this approach in order to compute the LOOCV error for a simple logistic regression model on the Weekly data set. Recall that in the context of classification problems, the LOOCV error is given in (5.4).

## (a) Fit a logistic regression model that predicts Direction using Lag1 and Lag2.

## (b) Fit a logistic regression model that predicts Direction using Lag1 and Lag2 ...
using all but the first observation.

## (c) Use the model from (b) to predict the direction of the first observation.
You can do this by predicting that the first observation will go up if P(Direction="Up"|Lag1, Lag2) > 0.5. Was this observation correctly classified?

## (d) Write a for loop...
from i = 1 to i = n, where n is the number of observations in the data set, that performs each of the following
steps:
<ol>
<li>Fit a logistic regression model using all but the ith observation to predict Direction using Lag1 and Lag2.</li>
<li>Compute the posterior probability of the market moving up for the ith observation.</li>
<li> Use the posterior probability for the ith observation in order to predict whether or not the market moves up.</li>
<li>Determine whether or not an error was made in predicting the direction for the ith observation. If an error was made, then indicate this as a 1, and otherwise indicate it as a 0.</li>
</ol>

## (e) Take the average of the n numbers obtained in (d4) ...
in order to obtain the LOOCV estimate for the test error. Comment on the results.


# Q8. 
We will now perform cross-validation on a simulated data set.

## (a) Generate a simulated data set as follows:

set .seed (1)

x=rnorm (100)

y=x-2* x^2+ rnorm (100)

In this data set, what is n and what is p? Write out the model used to generate the data in equation form.

## (b) Create a scatterplot of X against Y . Comment on what you find.


## (c) Set a random seed, and then compute the LOOCV errors...
that result from fitting the following four models using least squares:

i. Y = β0 + β1X + E

ii. Y = β0 + β1X + β2X2 + E

iii. Y = β0 + β1X + β2X2 + β3X3 + E

iv. Y = β0 + β1X + β2X2 + β3X3 + β4X4 + E.

Note you may find it helpful to use the data.frame() function to create a single data set containing both X and Y 
## (d) Repeat (c) using another random seed, and report your results.
Are your results the same as what you got in (c)? Why?

## (e) Which of the models in (c) had the smallest LOOCV error? 
Is this what you expected? Explain your answer.

## (f) Comment on the statistical significance ...
of the coefficient estimates that results from fitting each of the models in (c) using least squares. Do these results agree with the conclusions drawn based on the cross-validation results?