---
title: "CrimeLinkage"
author: "Diana McSpadden"
date: "2/17/2021"
output: html_document
---
# 1 Crime Linkage
Crime linkage attempts to determine if two or more unsolved crimes share a common offender. Pairwise crime linkage is the more simple task of deciding if two crimes were committed by a common offender; it can be considered a binary classification problem. The linkage data has 8 evidence variables that measure the similarity between a pair of crimes:

* spatial is the spatial distance between the crimes
* temporal is the fractional time (in days) between the crimes
* tod and dow are the differences in time of day and day of week between the crimes
* LOC, POA, and MOA are binary with a 1 corresponding to a match (type of property, point of entry, method of entry)
* TIMERANGE is the time between the earliest and latest possible times the crime could have occurred (because the victim was away from the house during the crime).
* The response variable indicates if the crimes are linked (y=1) or unlinked (y=0). The testing data does not include the response variable y.


```{r setup}
library(ggplot2)
library(tidyverse)
library(plotly)
library(dplyr)
library("GGally")
library(randomForest)
library(gridExtra)
library(broom)
library(caret)
library(kernlab)
```


#### Load the Data
```{r loaddata, echo=FALSE}
train = read_csv("linkage_train.csv")

test = read_csv("linkage_test.csv")
```

#### Set the Factors
```{r setFators}
train = train %>% 
  mutate(LOC = factor(LOC), POA = factor(POA), MOA = factor(MOA), y = factor(y)) 

test = test %>% 
  mutate(LOC = factor(LOC), POA = factor(POA), MOA = factor(MOA)) 

```

I want a subset of train to train with and a subset of train to test with.

```{r}
intrain <- createDataPartition(y = train$y, p= 0.7, list = FALSE)
trainTrain <- train[intrain,]
trainTest <- train[-intrain,]
```


```{r summarizeByCategory}
trainTrain %>%
  group_by(y) %>%
  summarize(N = n()) %>%
  mutate(Perc = round(N / sum(N), 2) * 100)
```

This is **very** unbalanced data.

#### View Pairwise
On the train, view pairwise correlations

```{r pairwise, fig.width=10, fig.height=10}
ggpairs(train, lower = list(continuous = "points", combo = "dot_no_facet"), progress = F)

```


```{r}
summary(train)
```

This is a very unbalanced data set. There are ways to create synthetic results, but I won't be using those here.


```{r setupTraining}
#using cross validation
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 10)
```

Center and scale because I don't know if temporal, tod, and dow are using the same scale for time differences.

```{r}
#crimeModel <- train(y~spatial+temporal+spatial:temporal+dow+tod+LOC+POA+MOA+I(dow^2)+I(tod^2)+I((temporal+spatial)^2, data = trainTrain, method = "glmnet", trControl=trctrl, tuneLength = 10)

crimeModel <- train(y~spatial+temporal+spatial:temporal+dow+tod+LOC+POA+MOA+I(dow^2)+I(tod^2)+I((temporal+spatial)^2), data = train, method = "glmnet", trControl=trctrl, tuneLength = 10)
```


```{r}
crimeModel
```

```{r}
#test_pred = predict(crimeModel, newdata = trainTest)

test_pred = predict(crimeModel, newdata = test, type="prob") %>% 
  mutate('y'=names(.)[apply(., 1, which.max)])
```

```{r}
# write to csv
```



```{r}
head(test_pred)
```


```{r}
#confusion matrices I used to evaluate the models I tried
confusionMatrix(table(test_pred, trainTest$y))
```

#### Results: THings I Tried and Didn't Use:

**all additive**
test_pred    0    1<br />
        0 8844   95<br />
        1    5   55<br />
        


**spatial+temporal+spatial:temporal+dow+tod+LOC+POA+MOA+I(temporal^2)+I(dow^2)+I(tod^2)+I((temporal+spatial)^2)+I((spatial+tod)^2)**
test_pred    0    1
        0 8844   87
        1    5   63
        
**spatial+temporal+spatial:temporal++dow+tod+LOC+POA+MOA+I(dow^2)++I(tod^2)**       
test_pred    0    1<br />
        0 8843   93<br />
        1    6   57<br />


**spatial+temporal+spatial:temporal+tod+POA+I(dow^2)**
test_pred    0    1<br />
        0 8848  110<br />
        1    1   40<br />

**spatial+temporal+spatial:temporal++dow+tod+LOC+POA+MOA+I(dow^2)++I(tod^2)**
test_pred    0    1<br />
        0 8848  110<br />
        1    1   40<br />



**spatial+temporal+spatial:temporal+tod+dow+LOC+POA+MOA+LOC:POA+LOC:MOA**
test_pred    0    1 <br />
        0 8843   94<br />
        1    6   56<br />
        
        
**spatial+temporal+spatial:temporal+spatial:tod+spatial:dow+tod+dow+LOC+POA+MOA+LOC:POA+LOC:MOA**
test_pred    0    1<br />
        0 8844   94<br />
        1    5   56<br />
        
        
#### Best Confusion Matrix I Produced:

**spatial+temporal+spatial:temporal+dow+tod+LOC+POA+MOA+I(dow^2)+I(tod^2)+I((temporal+spatial)^2**  
test_pred    0    1
        0 8845   87
        1    4   63




