---
title: "DataPrep-SandBox"
author: "Diana McSpadden"
date: "1/31/2021"
output: html_document
---
```{r}
library(tidyverse)

options(tibble.print_min = 5)

homes <- read_csv("https://raw.githubusercontent.com/uvastatlab/phdplus/master/data/albemarle_homes.csv")
```


```{r}
homes
```

```{r}
select(homes, totalvalue)
```
```{r}
select(homes, totalvalue, city)
```


```{r}
select(homes, 0:2) # select range of columns

```


```{r}
select(homes, -c(yearbuilt,condition)) #: select every column but

```

```{r}
select(homes, starts_with("f")) #: select columns that start with…


```

```{r}
select(homes, contains("f"))

```

```{r}
filter(homes, yearbuilt == 2006)
```

```{r}
filter(homes, bedroom == (max(bedroom) - 15) & yearbuilt < 1980)
```

```{r}
filter(homes, city %in% c("SCOTTSVILLE", "CROZET"))
```

## Boolean operators for multiple conditions

* a & b: and
* a | b: or
* xor(a,b): exactly or
* !a: not


```{r}
arrange(homes, finsqft)
```

```{r}
arrange(homes, desc(finsqft))
```

## Pipes!

The %>% is the pipe operator. It passes the result on left into the first argument of the function on the right, or “pipes” the previous output as the input for the next function. Read it in your head as then

For example, if we just want the totalvalue and lotsize for homes in Scottsville arranged in descending order of lotsize, without the pipe, we’d:

```{r}
scottsville <- filter(homes, city == "SCOTTSVILLE")
scottsville <- select(scottsville, totalvalue, lotsize)
scottsville <- arrange(scottsville, desc(lotsize))
scottsville
```


```{r}
homes %>% 
  filter(city == "SCOTTSVILLE") %>% 
  select(totalvalue, lotsize) %>% 
  arrange(desc(lotsize))
```
## Create new calculated columns
```{r}
homes %>% 
  mutate(value_sqft = totalvalue/finsqft) %>% 
  select(yearbuilt, condition, finsqft, totalvalue, city, value_sqft) %>% 
  arrange(desc(value_sqft), condition)
```

```{r}
homes %>% 
  mutate(value_sqft = totalvalue/finsqft,
         remodel = if_else(yearremodeled > 0, 1, 0)) %>% 
  select(value_sqft, remodel, city) %>% 
  arrange(value_sqft)
```

### mutate based on condition
if_else(condition, value_if_true, value_if_false, value_if_na)

### vectorize multiple if_else()
case_when(condition1 ~ value1, condition2 ~ value2, condition3 ~ value3, TRUE ~ value_everything_else)

## summarize()
Compute table of summaries.
```{r}
homes %>% 
  filter(yearbuilt > 0) %>% 
  summarize(oldest = min(yearbuilt), 
            newest = max(yearbuilt),
            avgyear = floor(mean(yearbuilt)),
            total = n())
```

### Summary functions include

* first(): first value
* last(): last value
* nth(.x, n): nth value
* n(): number of values
* n_distinct(): number of distinct values
* min(): minimum value
* max(): maximum value
* mean(): mean value
* median(): median value
* var(): variance
* sd(): standard deviation
* IQR(): interquartile range

## Group By
```{r}
homes %>% 
  filter(yearbuilt > 0) %>% 
  group_by(city) %>% 
  summarize(oldest = min(yearbuilt), 
            newest = max(yearbuilt), 
            avgyear = floor(mean(yearbuilt)),
            total = n())
```


## Other funtions

tally() - short hand for summarize(n())

count() - short hand for group_by() + tally()

summarize_all() - apply summary function to all variables

summarize_at() - apply summary function to selected variables

rename() - rename variables

recode() - modify values of variables


```{r}
summary(homes)
```

## Factors for categorical variables

```{r}
homes %>% count(condition) # currently a character
```

```{r}
homes %>% 
  mutate(condition = factor(condition)) %>% # make a factor
  count(condition)
```

** set factor with category labels in an order**

```{r}
cond_levels <- c("Excellent", "Good", "Average", "Fair", "Poor", "Substandard")
homes %>% 
  mutate(condition = factor(condition, levels = cond_levels)) %>% 
  count(condition)
```
### forcats

The forcats package, part of the tidyverse, provides helper functions for working with factors. Including

* fct_infreq(): reorder factor levels by frequency of levels
* fct_reorder(): reorder factor levels by another variable
* fct_relevel(): change order of factor levels by hand
* fct_recode(): change factor levels by hand
* fct_collapse(): collapse factor levels into defined groups
* fct_lump(): collapse least/most frequent levels of factor into “other”

## dplyr: Joins

Often, a single data set won’t contain all of the information or attributes necessary to address your questions. Instead, we have to join or combine multiple data sets.

**Mutating joins** use information from one data set to add variables to another data set (like mutate()).

**Filtering joins** use information from one data set to extract rows from another data set (like filter())

### mutating joins



* full_join(): keeps all observations in x and y
* left_join(): keeps all observations in x
* right_join(): keeps all observations in y
* inner_join(): keeps observations in both x and y
* filtering joins

## Filtering joins match observations in the same way, but rather than adding variables, it extracts observations.

* semi_join(): keeps all observations in x that have a match in y
* anti_join(): drops all observations in x that have a match in y



```{r}
?rm
```





