---
title: "DataViz-Sandbox"
author: "Diana McSpadden"
date: "1/31/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(scales)
library(ggplot2)

```

## The Grammar of Graphics
The Grammar of Graphics boiled down to 5 bullets, courtesy of Wickham (2016, p. 4):

* a statistical graphic is a mapping from data to aesthetic attributes (location, color, shape, size) of geometric objects (points, lines, bars).

* the geometric objects are drawn in a specific coordinate system.

* scales control the mapping from data to aesthetics and provide tools to read the plot (ie, axes and legends).

* the plot may also contain statistical transformations of the data (means, medians, bins of data, trend lines).

* faceting can be used to generate the same plot for different subsets of the data.

```{r}
github_url <- "https://github.com/uvastatlab/phdplus/raw/master/data/albemarle_homes.rds"
homes <- readRDS(url(github_url))
```

```{r}
head(homes)
```


```{r}
str(homes)
```

```{r}
homes <- homes %>% 
  mutate(isremodel = factor(remodel)) %>% # make a factor
  mutate(isfp = factor(fp)) %>% # make a factor
  mutate(landuse = factor(landuse)) %>% # make a factor
  mutate(issub = factor(issub))


str(homes)
```
```{r}
print(is.factor(homes$landuse))
homes$landuse <- factor(homes$landuse)
is.factor(homes$landuse)
```

```{r}
str(homes)
```


```{r}
homes$insub <- factor(homes$insub)
homes$remodel <- factor(homes$remodel)
homes$fp <- factor(homes$fp)

str(homes)
```

```{r}

vars <- c("CHARLOTTESVILLE", "CROZET", "EARLYSVILLE", 
          "KESWICK", "SCOTTSVILLE", "NORTH GARDEN")
homes <- homes %>% 
  filter(city %in% vars)
homes$city <- droplevels(homes$city) # droplevels resets the available factor values after subsetting a dataframe
```


```{r}
ggplot(homes, aes(x=finsqft, y=totalvalue, color=city)) + geom_point() 
```

```{r}
ggplot(filter(homes, city == "SCOTTSVILLE"), aes(x=finsqft, y=totalvalue, 
                 color=condition, size=bedroom)) + 
  geom_point()
```

```{r}
ggplot(homes, aes(x=finsqft, y=totalvalue, color=city)) + 
  geom_point() + 
  geom_smooth()
```


```{r}
ggplot(homes, aes(x=city, y=finsqft)) + 
  geom_boxplot()
```

## Moving beyond ggplot + geoms
A natural next step in exploratory graphing is to create plots of subsets of data. These are called **facets** in ggplot2.

Use facet_wrap() if you want to facet by one variable and have ggplot2 control the layout. Example:

+ facet_wrap( ~ var)

Use facet_grid() if you want to facet by one and/or two variables and control layout yourself.

Examples:
+ facet_grid(. ~ var1) - facets in columns
+ facet_grid(var1 ~ .) - facets in rows
+ facet_grid(var1 ~ var2) - facets in rows and columns

```{r}
ggplot(homes, aes(x=finsqft, y=totalvalue)) + 
  geom_point() + facet_wrap(~ city)
```


```{r}
ggplot(homes, aes(x=finsqft, y = stat(density))) +
  geom_histogram() + facet_grid(city ~ .) 
```


```{r}
```


```{r}
```





