---
title: "mod1_mcspadden-diana"
author: "Diana McSpadden (hdm5s)"
date: "2/2/2021"
output: html_document
---

# Module 1 Homework
## H. Diana McSpadden (hdm5s)

Complete exercises 1, 2, 6, 8, and 10 at the end of Chapter 2 of your textbook. Attempt each of these exercises prior to the live session since they will likely serve as material for discussion and interaction. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
```


## Exercise 1
For each of parts (a) through (d), indicate whether we would generally expect the performance of a flexible statistical learning method to be
better or worse than an inflexible method. Justify your answer.

#### (a) The sample size n is extremely large, and the number of predictors p is small.

ANSWER: **Inflexible**, this is an ideal regression data set. We can possibly create an accurate model that is also interpretable.

#### (b) The number of predictors p is extremely large, and the number of observations n is small.

ANSWER: **Inflexible** or **Flexible**, I could recommend two different approached: lasso method if the n is large enough and we want to select which of the parameters are most useful. Lasso will help with feature selection. However, if n is really small, a flexible method such a a support vector machine 

#### (c) The relationship between the predictors and response is highly non-linear.

ANSWER: **Flexible** because highly non-linear function decision boundaries require flexible approaches.

#### (d) The variance of the error terms, i.e. σ2 = Var(E), is extremely high.

ANSWER: This scenario requires a **flexible** method. Standard variance is an assumption for lasso, or regression methods and with a highly variable standard deviation

## Exercise 2
Explain whether each scenario is a classification or regression problem, and indicate whether we are most interested in inference or prediction.
Finally, provide n and p.

#### (a) We collect a set of data on the top 500 firms in the US...
For each firm we record profit, number of employees, industry and the CEO salary. We are interested in understanding which factors affect CEO salary.


ANSWER: This scenario is a **regression** problem specifically interested in determining **inference**, so an interpretable model is important. A regression solution will provide the coefficients for each parameter which allow comparison of parameter influence on CEO salary. n = 500, p = 3. 

#### (b) We are considering launching a new product and wish to know whether it will be a success or a failure...
We collect data on 20 similar products that were previously launched. For each product we have recorded whether it was a success or failure, price charged for the product, marketing budget, competition price, and ten other variables.


ANSWER: This is a **categorization** problem with "success" or "failure" categories. In this scenario we are interested in **predicting** whether our new product will succeed or fail. n = 20, p = 13.

#### (c) We are interest in predicting the % change in the USD/Euro

ANSWER: **Regression**, we are trying to **predict** a continuous value, not a condition. n = number of changes in USD/Euro in our data set. p = parameters in our data that we believe predict USD/Euro currency rate change.

## Exercise 6: 
Describe the differences between a parametric and a non-parametric statistical learning approach. What are the advantages of a parametric approach to regression or classification (as opposed to a nonparametric approach)? What are its disadvantages?


**ANSWER: **

Parametric approaches assume the shape of the function will be a specific linear shape. The advantage of a parametric approach is that it simplifies the fitting problem to estimating coefficients in the function, typically by reducing squared error from the mean, while keeping a linear function for the range. THe downside to assuming the form of the function is that the estimated function could be far from the true form. However, parametric approaches are 

Non-parametric approaches assume no such restrictions on the model's form, and thus try to map to the data points as closely as possible while minimizing variance (as configured). Non-parametric approaches require a large number of observations in order to limit over-fitting. Non-parametric approaches can also introduce excess bias, but if bias and over-fitting are controlled they may provide a better estimation of the funciton.


## Execise 8
This exercise relates to the College data set, which can be found in the file College.csv.

...

### (a) Use the read.csv() function to read the data into R. 
Call the loaded data college. Make sure that you have the directory set to the correct location for the data.


```{r q8a1}
college <- read.csv("College.csv")
head(college)
```

### (b) fix() ...
```{r q8b-1}
rownames(college) = college[,1] # first column is the name of the university
fix (college) #fix pops up a "fix" form to edit the data
```


```{r q8b-2}
college =college [,-1] #remove the first column
fix (college )
```

### (c) summary() and plotting

#### i: Use the summary() function
```{r q8ci-1}
summary(college) # display summary statistics and data information
```

#### ii. Use the pairs() function 
```{r q8cii-1, out.width = '80%'}
#?pairs
pairs(college[,2:11])
```

#### iii. plot() function
```{r q8ciii-1}
# plot side by side boxplots of Outstate versus Private (Private universities have more out of state students):
boxplot(college$Outstate~college$Private)
```

#### iv. Create a new qualitative variable, called Elite, ...
by binning the Top10perc variable. We are going to divide universities into two groups based on whether or not the proportion of students coming from the top 10% of their high school classes exceeds 50%.

```{r q8civ-1}
#?rep
Elite = rep("No",nrow(college )) # replicates the values in x.

Elite[college$Top10perc > 50]=" Yes" #for indexes where top 10% is above 50%, set Elite to "Yes"

Elite = as.factor (Elite) # set Elite as a factor Yes/No

college = data.frame(college ,Elite) # append the new Elite column to the college dataframe

```

```{r q8civ-2}
head(college) # show with new Elite column
```


```{r q8civ-3}
summary(college)
```

```{r q8civ-4}

boxplot(college$Outstate~college$Elite)
```

#### v. Use the hist() function to produce some histograms with differing numbers of bins for a few of the quantitative variables.





## Exercise 10