---
title: "mod9_mcspadden_diana"
author: "Diana McSpadden"
date: "4/13/2021"
output: html_document
---

# Module 9 Homework
## H. Diana McSpadden (hdm5s)

# Question 7 
In this problem, you will use support vector approaches in order to predict whether a given car gets high or low gas mileage based on the Auto data set.

```{r q7-setup}
library(ISLR)
library(tidyverse)
```

```{r q7-setup2}
auto = ISLR::Auto

head(auto)
```


## (a) Create a binary variable ...
that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.

**First:** I remove the name and the mpg columns.
```{r q7a-1}
#print(median(auto$mpg))
auto = auto %>% mutate(aboveMedian = ifelse(mpg > median(mpg), 1, 0) %>% factor()) %>% select(-name) %>% select(-mpg)

head(auto)
```


## (b) Fit a support vector classifier to the data ...
with various values of cost, in order to predict whether a car gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter. Comment on your results.

```{r q7b-setup}
library(e1071)
```

A linear kernel == a support vector classifier.

```{r q7b-1, cache=TRUE}
set.seed(1976)

svm_tuned = tune(svm, aboveMedian~., data=auto, kernel="linear", ranges=list(cost=c(0.001, 0.01, 0.1, 1)))

summary(svm_tuned)
```

**Answer Q7b**: The best cost was 1, with an error rate of ~0.087.

## (c) Now repeat (b), this time using SVMs with radial and polynomial basis kernels, 
with different values of gamma and degree and cost. Comment on your results.

First, I will try the radial models with different combinations of cost and gamma:
```{r q7c-1, cache=TRUE}
set.seed(1976)

svm_multi_tuned = tune(svm, aboveMedian~., data=auto, kernel="radial", ranges=list(cost=c(0.1,1,10,100),gamma=c(0.1,0.5,1,2)))

summary(svm_multi_tuned)


```
The best performing radial model had cost == 1, and gamma == 1, with an error rate of ~0.077.

```{r q7c-2, warning=FALSE, cache=TRUE}
set.seed(1976)

svm_poly_tuned = tune(svm, aboveMedian~., data=auto, kernel="polynomial", ranges=list(cost=c(0.1,1,10),gamma=c(0.5,1), degree=c(3,4)))

summary(svm_poly_tuned)
```
The best performing polynomial model had cost == 0.1, gamma == 1, degree of 3, with an error rate of ~0.082.

**Comments Q7c:** The radial model performed better than the polynomial model, leading me to believe that the relationship of the training data is NOT best represented with a linear or polynomial model, but needs the bell shaped curve of a radial model.

## (d) Make some plots to back up your assertions in (b) and (c).
Hint: In the lab, we used the plot() function for svm objects only in cases with p = 2. When p > 2, you can use the plot() function to create plots displaying pairs of variables at a time. Essentially, instead of typing

plot(svmfit , dat)

where svmfit contains your fitted model and dat is a data frame containing your data, you can type

plot(svmfit , dat , x1∼x4)

in order to plot just the first and fourth variables. However, you must replace x1 and x4 with the correct variable names. To find out more, type ?plot.svm.


```{r q7d-1}
?plot.svm

head(auto)

best_svm_model = svm(aboveMedian~., data=auto, type="C-classification", kernel="radial", cost=1, gamma=1)

par(mfrow=c(1,1))
#detach(auto)
plot(best_svm_model, auto, weight ~ horsepower, 
     slice = list(cylinders = median(auto$cylinders), displacement = median(auto$displacement), acceleration = median(auto$acceleration), year = mode(auto$year), origin = mode(auto$origin)))


```
```{r}
## more than two variables: fix 2 dimensions
data(iris)
head(iris)
m2 <- svm(Species~., data = iris)
plot(m2, iris, Petal.Width ~ Petal.Length,
     slice = list(Sepal.Width = 3, Sepal.Length = 4))


```


# QUestion 8
This problem involves the OJ data set which is part of the ISLR package.

## (a) Create a training set containing a random sample ...
of 800 observations, and a test set containing the remaining observations.

```{r q8a-1}

oj = ISLR::OJ

oj = oj %>% mutate(Purchase = factor(Purchase), StoreID = factor(StoreID), SpecialCH = factor(SpecialCH), SpecialMM = factor(SpecialMM), Store7 = factor(Store7), STORE = factor(STORE))

head(oj)

dim(oj)
```
```{r q8a-2}
set.seed(1976)
train = sample(nrow(oj), 870)

oj_train = oj[train,]
oj_test = oj[-train,]

```


## (b) Fit a support vector classifier ...
to the training data using cost=0.01, with Purchase as the response and the other variables as predictors. Use the summary() function to produce summary statistics, and describe the results obtained.
```{r q8b-1}
set.seed(1976)

q8b_svm = svm(Purchase~., data=oj_train, kernel="linear", cost=c(0.01))

summary(q8b_svm)

```

With what seems a relatively small cost (0.01), with a linear kernel, 468 observations of the 870 observations are support vectors. This seems high. 


## (c) What are the training and test error rates?

```{r q8c-1}
set.seed(1976)

q8b_tuned_svm = tune(svm, Purchase~., data=oj_train, kernel="linear", ranges=list(cost=c(0.01)))

summary(q8b_tuned_svm)

pred = predict(q8b_tuned_svm$best.model, newdata = oj_test)

length(pred)

table(true = oj_test$Purchase, pred)
```
The cross validation error on the training data is **0.1747**.

The error rate of the test data is: ((24 + 11) / 200) == **0.175**


## (d) Use the tune() function ...
to select an optimal cost. Consider values in the range 0.01 to 10.

```{r q8d-1}

set.seed(1976)
q8d_tuned_svm = tune(svm, Purchase~., data=oj_train, kernel="linear", ranges=list(cost=c(0.01,0.1,1,10)))

summary(q8d_tuned_svm)

```

**Answer Q8d:** The optimal cost with cross validation is **10**.

## (e) Compute the training and test error rates using this new value for cost.

```{r}
set.seed(1976)
q8e_tuned_svm = tune(svm, Purchase~., data=oj_train, kernel="linear", ranges=list(cost=c(0.01,0.1,1,10)))

summary(q8e_tuned_svm)

pred = predict(q8e_tuned_svm$best.model, newdata = oj_test)

length(pred)

table(true = oj_test$Purchase, pred)
```
**Answer Q8 e:** 

The best cost **training** error rate is **0.168**.

The **test** error rate using Cost = 10 is (23 + 12) / 200 == **0.175** - the same as the test error rate with cost == 0.01.

## (f) Repeat parts (b) through (e) using a support vector machine with a radial kernel. Use the default value for gamma.

```{r q8f-1}
print("Cost == 0.01")
set.seed(1976)
q8f_tuned_svm_01 = tune(svm, Purchase~., data=oj_train, kernel="radial", ranges=list(cost=c(0.01)))
summary(q8f_tuned_svm_01)
pred = predict(q8f_tuned_svm_01$best.model, newdata = oj_test)
length(pred)
table(true = oj_test$Purchase, pred)

print("Tuning COsts")
set.seed(1976)
q8f_tuned_svm_costs = tune(svm, Purchase~., data=oj_train, kernel="radial", ranges=list(cost=c(0.01,0.1,1,10)))
summary(q8f_tuned_svm_costs)
pred = predict(q8f_tuned_svm_costs$best.model, newdata = oj_test)
length(pred)
table(true = oj_test$Purchase, pred)


```
**Answer Q8f:**

**Using A Radial Kernel**

When using cost == **0.01**, the cross validation error rate on the training data is **0.39**, and the test error rate is 78 / 200 == **0.39**.


When tuning for cost, the best cost is **1**, and the training error rate is **0.168**, and the test error rate is (27 + 12) / 200 == **0.22**.

## (g) Repeat parts (b) through (e) using a support vector machine ...
with a polynomial kernel. Set degree=2.

```{r q8g-1}
#set.seed(1976)

#svm_poly_tuned = tune(svm, aboveMedian~., data=auto, kernel="polynomial", ranges=list(cost=c(0.1,1,10),gamma=c(0.5,1), degree=c(3,4)))


print("Cost == 0.01")
set.seed(1976)
q8g_tuned_svm_01 = tune(svm, Purchase~., data=oj_train, kernel="polynomial", ranges=list(cost=c(0.01), degree=c(2)))
summary(q8g_tuned_svm_01)
pred = predict(q8g_tuned_svm_01$best.model, newdata = oj_test)
length(pred)
table(true = oj_test$Purchase, pred)

print("Tuning COsts")
set.seed(1976)
q8g_tuned_svm_costs = tune(svm, Purchase~., data=oj_train, kernel="polynomial", ranges=list(cost=c(0.01,0.1,1,10), degree=c(2)))
summary(q8g_tuned_svm_costs)
pred = predict(q8g_tuned_svm_costs$best.model, newdata = oj_test)
length(pred)
table(true = oj_test$Purchase, pred)


```
**Answer Q8g:**

**Using A Polynomial Kernel of Degree 2**

When using cost == **0.01**, the cross validation error rate on the training data is **0.39**, and the test error rate is 78 / 200 == **0.39**.


When tuning for cost, the best cost is **10**, and the training error rate is **0.18**, and the test error rate is (29 + 12) / 200 == **0.205**.


## (h) Overall, which approach ...
seems to give the best results on this data?

**Answer Q8h:**
The linear, support vector classifier with a Cost == 1 provides the best performance on the test data set. The true relationship is linearly separable at the higher SV dimension.

